---
title: "Day04-Notes"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


## Bivariate Regression

```{r}
library(MASS)
set.seed(1234)
varX = 100
varY = 400

## covariance = how both variables are varying together
covXY = 100
covmatrix <- matrix(c(varX, covXY, covXY, varY), 2, 2)

## 40 is the mean of X, 200 is the mean of Y
population <- as.data.frame(mvrnorm(n = 1000, 
                                     mu = c(40, 200), 
                                     Sigma = covmatrix, 
                                     empirical = TRUE))
names(population) = c("X", "Y")
```

### Plot
```{r}
library(ggplot2)
ggplot(population, aes(X, Y)) +
  geom_point()
```

### Line of best fit
```{r}
poplm <- lm(Y ~ X, data = population)
```

### Let's take a sample and see what our relationship is...

```{r}
library(dplyr)
samplesize <- 20
sample.data <- population %>% sample_n(samplesize)
lm(Y ~ X, data = sample.data)
```

```{r}
library(broom)
modeloutput <- tidy(lm(Y ~ X, data = sample.data))

for(i in 2:100) modeloutput <- lm(Y ~ X, data = population %>% sample_n(samplesize)) %>% tidy %>% bind_rows(modeloutput)
```

```{r}
modeloutput %>% filter(term == "X") %>% ggplot(aes(estimate)) +
  geom_histogram()
```

```{r}
library(stargazer)
modeloutput %>% filter(term == "X") %>% stargazer(type = "text")
```

## New Data Generating Process

```{r}
varX = 100
varerror = 400

## if there is covariance between error and X, sample coefficient cannot be a good estimate of relationship.
covXerror = 0
covmatrix2 <- matrix(c(varX, covXerror, covXerror, varerror), 2, 2)

## mean of error = 0
Xe <- as.data.frame(mvrnorm(n = 1000, 
                                     mu = c(40, 0), 
                                     Sigma = covmatrix2, 
                                     empirical = TRUE))
names(Xe) = c("X", "error")

beta0 <- 120
beta1 <- 1

## make a Y variable with Y = beta0 + beta1*X + error
Xe <-  Xe %>% mutate(Y = beta0 + beta1*X + error)
```

### Plot
```{r}
library(ggplot2)
ggplot(Xe, aes(X, Y)) +
  geom_point()
```

### Line of best fit
```{r}
newlm <- lm(Y ~ X, data = Xe)
```

### Let's take a sample and see what our relationship is...

```{r}
samplesize2 <- 1000
sample.data2 <- Xe %>% sample_n(samplesize2)
lm(Y ~ X, data = sample.data2)
```

```{r}
modeloutput2 <- tidy(lm(Y ~ X, data = sample.data2))

for(i in 2:1000) modeloutput2 <- lm(Y ~ X, data = Xe %>% sample_n(samplesize)) %>% tidy %>% bind_rows(modeloutput2)
```

```{r}
modeloutput2 %>% filter(term == "X") %>% ggplot(aes(estimate)) +
  geom_histogram()
```

```{r}
modeloutput2 %>% filter(term == "X") %>% stargazer(type = "text")
```

### How can we move the distribution either to the right or to the left?

**1.** Change the Covariance (make an error no longer a "white noise")
**2.** Just change the true mean

### How can we change the spread of the ditribution

**1.** Change the sample size(n)
**2.** Change the variance of population: variance of error and X