---
title: "Day03-Notes"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## Important Statistical Distributions

* z (normal)
* t (almost normal)
* $\chi^2$ (sum of squared normal)

### Central Limit Theorem

If we know the population mean and variance... then we predict that the sample mean will be normally distributed.

If the population has a mean of 144 and a variance of 36, what will the sample mean distribution tend to look like if the sample size is 100 and we took 400 samples?

```{r}
library(MASS)
set.seed(1239876)
population <- data.frame(X = mvrnorm(n = 10000, 
                                     mu = 144, 
                                     Sigma = 36, 
                                     empirical = TRUE))
```

```{r}
library(stargazer)
stargazer(population, type = "text")
```

```{r}
library(ggplot2)
ggplot(population, aes(X)) + geom_histogram(binwidth = 2)
```


Let's run a simulation...

```{r}
library(dplyr)
samplesize = 25
mysample <- sample_n(population, samplesize, replace = TRUE) 
mysample %>% summarize(mean(X))
mysample %>% summarize(sd(X))
```

```{r}
num_sims <- 100

sim.out <- population %>% sample_n(samplesize) %>% 
  summarise(xbar = mean(X), sdhat = sd(X))

for (i in 2:num_sims) {
sim.out <- population %>% sample_n(samplesize) %>% 
  summarise(xbar = mean(X), sdhat = sd(X)) %>% bind_rows(sim.out)
}
```

```{r}
stargazer(sim.out, type = "text")
```

```{r}
sim.out %>% ggplot(aes(xbar)) + 
  geom_histogram(aes(y = ..density..), bins = 20) +
  stat_function(fun = dnorm, 
                args = list(mean = 144, sd = 6/sqrt(samplesize)))
``` 


## Hypothesis Testing

Imagine testing the hypothesis that the true population mean was 144. How would we conduct this hypothesis test? 

$$\frac{\bar{x} - \mu}{\widehat{\sigma_\bar{x}}} = ?$$

```{r}
sim.out <- sim.out %>% 
  mutate(tstat = (xbar - 144)/(sdhat/sqrt(samplesize)))

sim.out %>% ggplot(aes(tstat)) + 
  geom_histogram(aes(y = ..density..), bins = 20) +
  stat_function(fun = dt, 
                args = list(df = samplesize - 1))
```


## Confidence Intervals

```{r}
my.alpha <- 0.05
test.df <- samplesize - 1
tstat.CI <- qt(my.alpha/2, test.df)

sim.out <- sim.out %>% 
  mutate(run = row_number(),
         low = xbar + tstat.CI*sdhat/sqrt(samplesize),
         high = xbar - tstat.CI*sdhat/sqrt(samplesize))

ggplot(sim.out, aes(run, xbar)) + 
  geom_point() + 
  geom_segment(aes(x = run, y = low, 
                   xend = run, yend = high))
```

**CI of each sample is different because standard error of each sample varies (sdhat/sqrt(samplesize)).**

What if we want to make the CIs that don't include the true value a different color?

```{r}
sim.out <- sim.out %>% 
  mutate(Includes = if_else(144 > low & 144 < high, "yes", "no"))

ggplot(sim.out, aes(run, xbar)) + 
  geom_point() + 
  geom_hline(yintercept = 144) +
  geom_segment(aes(x = run, y = low, 
                   xend = run, yend = high, color = Includes))
```


