---
title: "Day09-Notes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Day08-data

```{r}
library(readxl)
day08 <- read_excel("~/Documents/2017-18/ECON329/notes/Day09-data.xlsx")
day08_log.lm <- lm(log(Y) ~ log(X2) + log(X3), data = day08)
summary(day08_log.lm)
```

```{r}
F <- (0.9642/2)/((1-0.9642)/48)
p_value <- 1-pf(F,2,48)
```


```{r}
day08_log.lm2 <- lm(log(Y) ~ log(X2), data = day08)
summary(day08_log.lm2)
```

F-stat = t-stat^2 (when it comes to one variable SLR)

## Is coefficients are equal to each other

t-stat = $\frac{(\widehat{\beta} - \widehat{\gamma}) - (\beta - \gamma)}{se(\widehat{\beta} - \widehat{\gamma})}$
se($\widehat{\beta} - \widehat{\gamma}$) = $\sqrt{var(\widehat{\beta}) + var(\widehat{\gamma}) - 2cov(\widehat{\beta},\widehat{\gamma})}$

```{r}
obs_sum <- day08_log.lm$coefficients[2] - day08_log.lm$coefficients[3]

var_x2 <- vcov(day08_log.lm)["log(X2)","log(X2)"]
var_x3 <- vcov(day08_log.lm)["log(X3)","log(X3)"]
cov_x2_x3 <- vcov(day08_log.lm)["log(X2)","log(X2)"]
t_stat <- obs_sum / sqrt(var_x2+var_x3-2*cov_x2_x3)
```

## Is coefficients sum up to 1

$\beta + \gamma$ = 1

t-stat = $\frac{(\widehat{\beta} + \widehat{\gamma}) - (\beta + \gamma)}{se(\widehat{\beta} + \widehat{\gamma})}$
se($\widehat{\beta} + \widehat{\gamma}$) = $\sqrt{var(\widehat{\beta}) + var(\widehat{\gamma}) + 2cov(\widehat{\beta},\widehat{\gamma})}$


```{r}
obs_sum_2 <- coef(day08_log.lm)["log(X2)"] + coef(day08_log.lm)["log(X3)"]

VCOV <- vcov(day08_log.lm)

seBetas <- sqrt(VCOV["log(X2)", "log(X2)"] + 
  VCOV["log(X3)", "log(X3)"] +
  2*VCOV["log(X2)", "log(X3)"])

(obs_sum_2 - 1) / seBetas
```

## Residual Sum

```{r}
sum(day08_log.lm$residuals^2)
```

