---
title: "Day15-Notes"
author: "Colin Pi"
date: "5/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pooled OLS

```{r}
table16 <- read.csv("~/Documents/2017-18/ECON329/notes/Day15-files/Table16.1.csv")
pooled.lm <- lm(C ~ Q + PF + LF, data = table16)
summary(pooled.lm)
```

## Fixed Effect (with different intercept based on airline)

```{r}
fixed.lm1 <- lm(C ~ Q + PF + LF + factor(I), data = table16)
summary(fixed.lm1)
```

## Fixed Effect (with different intercept based on airlines + time effect)

```{r}
fixed.lm2 <- lm(C ~ Q + PF + LF + factor(I) + T, data = table16)
summary(fixed.lm2)
```

## Fixed Effect (with different intercept based on time)

```{r}
fixed.lm3 <- lm(C ~ Q + PF + LF + factor(T), data = table16)
summary(fixed.lm3)
```

## Fixed Effect (with different intercept based on time and airline)

```{r}
fixed.lm4 <- lm(C ~ Q + PF + LF + factor(I) + factor(T), data = table16)
summary(fixed.lm4)
```

## Fixed Effect (with different intercept and effect of Q based airline)

```{r}
fixed.lm5 <- lm(C ~ Q*factor(I) + PF + LF, data = table16)
summary(fixed.lm5)
```

## Use plm the annoying package

```{r}
library(plm)
pooled.lm2 <- plm(C ~ Q + PF + LF, data = table16, model = "pooling", index = c("I","T"))
summary(pooled.lm2)
```

## Within effect

```{r}
within.lm <- plm(C ~ Q + PF + LF, data = table16, model = "within", index = c("I","T"))
summary(within.lm)
```

## Random effect!

```{r}
random.lm <- plm(C ~ Q + PF + LF, data = table16, model = "random", index = c("I","T"))
summary(random.lm)
```

## Hausman Test

```{r}
phtest(random.lm, within.lm)
```

Reject the null hypothesis that both models are the same (meaning that fixed effects model is needed)

## Breush - Pagan Lagrange Multiplier Test

```{r}
plmtest(pooled.lm2, type = c("bp"))
```

